---
title: "Clustering analysis"
author: "Yifei Wan"
date: "November 27, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
#1 .Input data
library(factoextra)
#setwd('c:/users/yifei.wan/Desktop')
count_genus_raw <- read.csv('Genus_Level_Aggregate_Counts.csv', header = T, row.names = 1)
count_genus_t <- t(count_genus_raw) ## transpose matrix
```

# 1. Normalize data

The raw count would be normalized by t-statistic and feature scaling method respectively as `count_nor.z` and `count_nor.s`.

```{r, echo=F}
count_nor.z <- scale(count_genus_t)
count_nor.s <- apply(count_genus_t, 2, function(count){(count - min(count, na.rm = T)) / (max(count, na.rm = T) - min(count, na.rm = T))})
```


# 2. Assessing Clustering Tendency

## 2.1 Hopkins statistic
Apply Hopkins statistic to evaluate the clustering tendency of count data. A value close to 0 tends to indicate the data is highly clustered, while uniformly randomly distributed points will tend to result in values around 0.5.

```{r, echo=F}
hopkins_stat.z <- get_clust_tendency(count_nor.z, n = nrow(count_nor.z)-1, graph = F)
hopkins_stat.s <- get_clust_tendency(count_nor.s, n = nrow(count_nor.s)-1, graph = F)
```

+ Hopkins statistic based on t-statistic normalization
```{r, echo=F}
hopkins_stat.z$hopkins_stat
```

+ Hopkins statistic based on feature scaling normalization
```{r, echo=F}
hopkins_stat.s$hopkins_stat
```

Two values are close to 0.5 which implies that data points (samples) uniformly randomly distributed somehow. They may not be clusterable. 

## 2.2 Dissimilarity matrix

Visualize the distance matrix to review the dissimilarity within dataset. If data points are clusterable, the matrix would display somewhat clutering pattern.

+ Dissimilarity matrix based on t-statistic normalization

Low value indicate high similarity.


```{r, echo = F}
fviz_dist(dist(count_nor.z), order = TRUE, show_labels = TRUE, lab_size = NULL,
        gradient = list(low = "red", high = "white"))
```




+ Dissimilarity matrix based on feature scaling normalization

```{r, echo=F}
fviz_dist(dist(count_nor.s), order = TRUE, show_labels = TRUE, lab_size = NULL,
          gradient = list(low = "red", high = "white"))
```

Both images don't appear any remarkable clustering patterns.


# 3. PCA analysis

Apply PCA to reduce the dimensions and analyze pattern of dataset.

+ PCA based on t-statistic normalization

```{r, echo=F}
fviz_pca_ind(prcomp(count_nor.z), title = "PCA : Genus-level with t-statistic normaliztion", 
             palette = "jco", habillage = rep(c('Treatment', 'Control'), 5),
             geom = "point", ggtheme = theme_classic(),
             legend = "bottom", label='all')
```



+ PCA based on feature scaling normalization
```{r, echo=F}
fviz_pca_ind(prcomp(count_nor.s), title = "PCA : Genus-level with t-statistic normaliztion", 
             palette = "jco", habillage = rep(c('Treatment', 'Control'), 5),
             geom = "point", ggtheme = theme_classic(),
             legend = "bottom", label='all')
```

About PCA result implies that there is no linear boundary between two groups.


# 4. Clustering

```{r, echo=F}
cluster.z <- kmeans(count_nor.z, centers = 2, iter.max = 1e6)
cluster.s <- kmeans(count_nor.s, centers = 2, iter.max = 1e6)
fviz_cluster(list(data = count_nor.z, cluster = cluster.z$cluster),habillage = rep(c('Treatment', 'Control'), 
             ellipse.type = "norm", geom = "point", stand = FALSE, palette = "jco", ggtheme = theme_classic(), label = 'all'))
fviz_cluster(list(data = count_nor.s, cluster = cluster.s$cluster),habillage = rep(c('Treatment', 'Control'), 
             ellipse.type = "norm", geom = "point", stand = FALSE, palette = "jco", ggtheme = theme_classic(), label = 'all'))             
```


# 5. Conclusion

The genus level count data is not clusterable.